\chapter{Acknowledgments}
A Ph.D. is not easy and there's a long list of people to whom I owe a debt of gratitude. This is an incomplete list, but I will mention as many as I can.
First, my advisors and mentors through grad school, Alexey and Brian, were paragons of unquenchable curiosity and intellectual voracity. They were also extremely generous and patient, and always there to lend an ear. I thank them both for their guidance and support.

I also want to thank the postdocs with whom I've had the pleasure of learning from, including Zhe-Xuan Gong, Mike Foss-Feig, and Shenglong Xu, former graduate students Zachary Eldredge and Jeremy Young, and later my Flatiron mentors Peter Lunts, Miles Stoudenmire, and Matt Fishman. I thank my collaborators on everything Lieb-Robinson---Minh Tran, Adam Ehrenberg, and Abhinav Deshpande---and on scrambling---Tianci Zhou and Xiao Chen.

A list of postdoc mentors who lent their advice:
Chris Baldwin,
Jim Garrison,
Lucas Brady,
Luis-Pedro Garcia,
Przemek Bienias,
Simon Lieu, and
Zhicheng Yang.
A list of QuICS/JQI and CS graduate students:
Alireza Seif,
Aniruddha Bapat,
Denis Peskov,
Dhruv Devulapalli,
Eddie Schoute,
Fangli Liu,
Jon Kunjummen,
Jonathan Curtis,
Manasi Shingane,
Matt Kovacs-Deak,
Naren Manjunath,
Nhung Nguyen,
Pradeep Niroula,
Ron Belyansky,
Su-Kuan Chu,
Subhayan Sahu,
Yidan Wang,
Yuan Su, and
Yusuf Alnawaktha.

Thanks to the staff of QuICS and JQI. In particular, thanks to Javiera Caceres, Andrea Svejda, Melissa Britton for helping with travel and finances, and Josiland for helping me through the graduation paperwork.
I want to thank members of the GRAD-MAP team for working to support diversity, especially my co-leads, Milena Crnogorcevic and Charlotte Ward. Thanks to various members of the Physics and CS departments for the fun that helped preserve sanity, including board games and volleyball/frisbee, to the members of the Linke Lab who adopted me and took me along camping and beach trips, and finally to the Klein's Group of (approximately) Four: Abhinav, Ani, Eddie, and Minh, for being solid friends through thick and thin.

Along the road to this Ph.D. were the fair share of stumbles and failures. The true credit goes out to those who stayed by my side and stepped up when times were roughest. To that end, I thank my family---my mom, dad, grandma, and Vivian, as well as my extended family---for always having my back, Alexey for his patience and understanding and, of course, my beloved Yingyue Zhu, whose love and devotion has been a steadfast pillar of support throughout the good times and the bad. Thank you.
